Errorless irrationality: removing prediction errors from the inverse base-rate effect

### 2022-12-19 docs/manuscript.tex
continue related work

We will build on work carried out by \citeA{johansen2007paradoxical}.
\citeA{johansen2007paradoxical} failed to observe the inverse base-rate effect in an observational learning condition, where category labels were presented with features at the same time to participants.
They largely interpreted their results in opposition to EXIT \cite{kruschke2001toward}, but their experimental procedure does not correspond to the standard inverse base-rate effect procedure.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared features during training that facilitates attentional relocation, which in turn moves responding towards the irrational rare preference.
The only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
A simple explanation for that result is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.

@@

Below, we detail our first attempt to test whether we could observe the rare response bias without any explicit error-driven psychological mechanism.

@@

- auxiliary phenomenon (Wills CIRP addition)
- Things that any theory of IBRE should explain
- current theories fall short

>>>>

We will build on work carried out by \citeA{johansen2007paradoxical}.
\citeA{johansen2007paradoxical} failed to observe the inverse base-rate effect in an observational learning condition, where category labels were presented with features at the same time to participants.
They largely interpreted their results in opposition to EXIT \cite{kruschke2001toward}, but their experimental procedure does not correspond to the standard inverse base-rate effect procedure.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared features during training that facilitates attentional relocation, which in turn moves responding towards the irrational rare preference.
The only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
A simple explanation for that result is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.
Their design is therefore not sufficient to answer our question, whether prediction error is a necessary condition.

@@

Below, we detail our first attempt to test whether we could observe the rare response bias without an explicit error-driven psychological mechanism.
The design component which is most likely to result in any error-driven tuning is feedback.
To remove feedback, we will present category labels with their respective features.

@@

- auxiliary phenomenon (Wills CIRP addition)
- Things that any theory of IBRE should explain
- current theories fall short
- eye-tracking and attention
